### 2025-12-11 reviewer_response_todo.md

# Reviewer Response Completeness Check & TODO List

## ✅ ADDRESSED ITEMS

### Referee #1 Minor Details:
- ✅ Comment 4: Predictive medical research tradition - Steyerberg reference added (line 381)
- ✅ Comment 5: Interpretability and risk metrics - Addressed (lines 390-405)
- ✅ Comment 6: Temporal modeling vs age-dependent risk - Addressed (lines 407-428)
- ✅ Comment 7: Heritability estimates - Addressed with CVD comparison (lines 457-483)
- ✅ Comment 8: Joint phenotype modeling - Addressed (lines 485-490)
- ✅ Comment 9: AUC comparisons - Comprehensive comparisons added (lines 491-628)
- ✅ Comment 10: Age-specific discrimination - Clarified (lines 630-677)
- ✅ Comment 11: Figure organization - Addressed (lines 678-683)

### Referee #2 Minor Concerns:
- ✅ Cohort Definition - Detailed in Methods (lines 900-906)
- ✅ Genetic Analysis - Detailed ancestry handling (lines 907-912)
- ✅ Phenotype Handling - Comprehensive ICD/PheCode transformation details (lines 914-927)
- ✅ Analytical Decisions - Justification for 348 diseases and K=20 (lines 929-933)
- ✅ Code Availability - Fixed GitHub link (line 935)

### Referee #3 Minor Comments:
- ✅ Comment 8: Heterogeneity terminology - Clarified (line 1270)
- ✅ Comment 9: Signature count (20 vs 21) - Confirmed K=20 (line 1272)
- ✅ Comment 10: Fig 2B x-axis - Clarified (line 1274)
- ✅ Comment 11: Fig 4D clustering - Addressed (line 1276)
- ✅ Comment 12: SBayesS/LDpred - Acknowledged (line 1278)
- ✅ Comment 13: geq1000 notation - Corrected to ≥1,000 (line 1280)
- ✅ Comment 14: Fig 4B clusters - Clarified methodology (line 1282)
- ✅ Comment 15: Harrell's C - Explained why AUC is appropriate (line 1284)
- ✅ Comment 16: AUC invariance - Acknowledged limitation (line 1286)
- ✅ Comment 17: Computational intensity - Addressed (line 1288)

---

## ❌ TODO ITEMS

### 1. **Missing Reference: "Stires and Briggs"**
   - **Location**: Reviewer comment line 55 mentions "Stires and Briggs" as a book reference
   - **Current Status**: Only Steyerberg (2019) is cited in response (line 381)
   - **Action Needed**: 
     - Verify if "Stires and Briggs" is a real reference (may be a typo or alternative name)
     - If valid, add to the response alongside Steyerberg
     - If not found, confirm Steyerberg alone is sufficient (it's a well-known comprehensive text)

### 2. **Computational Hours Placeholder**
   - **Location**: Line 1288 in reviewer_response.tex
   - **Current Text**: "full UK Biobank analysis (39 subsets) required ~[X] compute hours"
   - **Action Needed**: 
     - Fill in actual compute hours or provide estimate
     - If exact number unavailable, provide reasonable estimate based on:
       - 200 epochs per subset
       - Number of subsets (39)
       - Approximate time per epoch
     - Alternative: State "approximately X hours" or provide range

### 3. **Simulation Study Mention**
   - **Location**: Simulation study exists in current.tex (lines 495-503) but not explicitly mentioned in reviewer response
   - **Current Status**: Simulation study validates model recovery but not highlighted in response
   - **Action Needed**: 
     - Consider adding brief mention in "Model Robustness and Validity" section
     - Or add to "Model Validity and Learning Dynamics" (Analysis 21)
     - Note: Not explicitly requested by reviewers, but demonstrates model validation

---

## 📊 ASSESSMENT: Code Detail & Simulation Comprehensiveness

### Code Detail: ✅ **COMPREHENSIVE**
- ✅ GitHub repository link fixed and accessible
- ✅ All 26 analysis notebooks documented
- ✅ Code structure explained (PyTorch implementation)
- ✅ Training scripts, prediction pipelines documented
- ✅ Vectorized implementation mentioned
- ✅ Computational details provided (epochs, convergence, inference time)
- ✅ Data availability commitments detailed

### Simulation Study: ✅ **ADEQUATE BUT COULD BE ENHANCED**

**Current Simulation (current.tex lines 495-503):**
- ✅ Uses model itself as generative model (good validation approach)
- ✅ Tests recovery of: cluster structure, temporal dynamics, genetic effects
- ✅ Reports: correct cluster count (5/5), Jaccard similarity (0.795)
- ✅ Demonstrates model can recover known parameters

**Potential Enhancements (not critical but would strengthen):**
- Could add: Sensitivity analysis (varying N, D, K)
- Could add: Performance under different noise levels
- Could add: Comparison with alternative initialization strategies
- **Note**: Current simulation is sufficient for validation purposes

---

## 🎯 PRIORITY ACTIONS

**HIGH PRIORITY:**
1. Fill in computational hours placeholder [X] → actual number/estimate
2. Verify/address "Stires and Briggs" reference

**LOW PRIORITY (Optional):**
3. Add brief mention of simulation study to response (if space allows)

---

## 📝 NOTES

- Most minor details are comprehensively addressed
- Methods section in current.tex appears thorough
- Simulation study validates model appropriately
- Main gaps are: computational hours number and potential missing reference

### 2025-12-18 rebuttal_consistency_response.tex

% Response to Referee \#1, Question 8 (and Referee \#2 interpretability concerns)

We acknowledge the reviewer's concern that modeling correlated phenotypes jointly can reduce specificity and potentially compromise interpretability. However, our analyses demonstrate that the disease-to-signature assignments in \aladyn{} are highly robust and interpretable.

To assess the stability and interpretability of signature assignments, we evaluated consistency across independent data subsets. We split the UK Biobank data into 40 independent batches (each containing 10,000 individuals), fit the model separately to each batch, and examined whether each disease's top signature (the signature with the highest PSI value for that disease) remained consistent across batches.

\textbf{Results:} For all 348 diseases, the top signature assignment was identical across all 40 batches (100\% consistency). This finding demonstrates that disease-to-signature assignments are stable and reproducible across independent data subsets, rather than being unstable amalgamations that vary with data sampling.

This high consistency across 40 independent batches provides several important insights:
\begin{itemize}
    \item \textbf{Interpretability:} The stable assignment of diseases to signatures across 40 independent data subsets indicates that the signatures capture reproducible, interpretable disease relationships rather than arbitrary data-driven patterns.
    \item \textbf{Robustness:} The consistency across 40 batches demonstrates that signature structure is not dependent on specific subsets of individuals, supporting the generalizability of our findings.
    \item \textbf{Biological coherence:} The fact that each disease consistently associates most strongly with the same signature across all 40 batches suggests that signatures reflect coherent biological processes, which enhances their interpretability despite being composites of multiple diseases.
\end{itemize}

While we acknowledge that signature-based associations may be less specific than individual disease associations, the stability demonstrated here indicates that the loss of specificity comes with a gain in robustness and interpretability. The signatures represent stable, reproducible patterns of disease co-occurrence that can be consistently identified across independent data subsets, making them suitable for both discovery (as evidenced by the novel genetic associations reported) and clinical application (as evidenced by the predictive performance improvements).
